---
title: "`TrainPredict` Package"
output: github_document
---


## Overview

`TrainPredict` is an R package designed to streamline the process of logistic regression modeling for binary classification tasks. With functions for splitting data, training models, making predictions, and assessing model performance.

## Installation

To install the package directly from GitHub, use the following command:

```{r, eval=F}
# Install from GitHub using devtools
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
devtools::install_github("AU-R-Programming/Final_Project_Group_7")
```

## Getting Started

Once installed, load the `TrainPredict` package to start using its functionality:

```{r}
library(TrainPredict)
```

## Example Workflow

Here is a step-by-step workflow that demonstrates the main functions of the `TrainPredict` package using the built-in `mtcars` dataset.


### 1. Split Dataset into Training and Test Sets

The `train_test_sampling()` function is used to split the dataset into training and test sets. Here, we use an 75-25 split.

```{r}
# Split the dataset into training and test sets (75-25 split).
# We will use variable `am` as the dependent binary variable. 
data("mtcars")
split_data <- train_test_sampling(mtcars, dependent_var = "am", train_prop = 0.75, return_data = TRUE, seed = 123)

# Extract training and test sets.
train_data <- split_data$train
head(train_data)

test_data <- split_data$test
head(test_data)

dim(mtcars)
dim(train_data)
dim(test_data)
```

### 2. Train a Logistic Regression Model

We use the `lr()` function to train a logistic regression model using selected predictors.

```{r}
# Train a logistic regression model using 'hp', 'mpg', and 'wt' as predictors of the 'am' binary variable.
model <- lr(am ~ hp + mpg + wt, data = train_data, B=100, alpha=0.05)

# Display model coefficients.
print(model$beta_optimized)

# Display confusion matrix.
print(model$confusion_matrix)
```

### 3. Make Predictions on Test Data

Evaluate the model's performance on the test dataset using the `predict_test()` function.

```{r}
# Make predictions on the test dataset
test_predictions <- predict_test(model = model, new_data = test_data, dependent_variable_col = "am")

# Display predictions
(test_predictions)
```

### 4. Make Predictions on New Data

Use the `predict_new()` function to make predictions for new data points.

```{r}
# Create new data points for prediction
new_data <- data.frame(hp = c(225, 110),
                       mpg = c(15.9, 25.6),
                       wt = c(3.485, 1.985))

# Predict using the trained model
new_predictions <- predict_new(data = new_data, model = model, threshold = 0.5)

# Display predictions
print(new_predictions)
```

## Key Functions

- `train_test_sampling()`: Split your dataset into training and test sets.
- `lr()`: Train a logistic regression model.
- `predict_test()`: Make predictions on the test dataset.
- `predict_new()`: Make predictions on new data using a model trained with `lr()`.

## Conclusion

The `TrainPredict` package makes logistic regression modeling easy and efficient by providing key functionalities for data splitting, model training, and prediction. Whether you are a beginner in machine learning or an experienced data scientist, `TrainPredict` can help you simplify the modeling workflow.

## Get Involved

If you'd like to contribute or report any issues, please visit the GitHub repository:

[TrainPredict on GitHub](https://github.com/AU-R-Programming/Final_Project_Group_7)


